---
title: "ParlGov Visualizations -- Elections"
output:
  html_document:
    code_folding: hide
    fig_width: 8 
    fig_height: 5 
    theme: cosmo 
---

```{r package, include = FALSE}
library(tidyverse)
library(countrycode)
library(grid)
library(ggrepel)
library(lubridate)
```

```{r datasets, include = FALSE}
election_raw <- read_csv("https://parlgov.org/data/parlgov-development_csv-utf-8/view_election.csv", 
                          locale = locale(encoding = "utf-8"))
lces <- read_csv("data/vdem_lces_new.csv") # for further differentiation among electoral systems
```

```{r data wrangling and precaution, include = FALSE}
## Election ----
election <- election_raw %>% 
  mutate(vote_share_0 = vote_share / 100,
                year = as.numeric(substr(election_date, 1, 4))) %>% 
  group_by(election_id, year) %>% 
  filter(!is.na(vote_share_0)) %>% 
  mutate(enep = 1 / sum(vote_share_0^2), # ENEP
         enpp = 1 / sum((seats/seats_total)^2), # ENPP
         region = as.factor(countrycode(country_name, "country.name", "region23")) # region
         ) 

election_enp <- election %>% 
  select(country_name, country_name_short, region, election_type, year, election_id, election_date, enep, enpp) %>% 
  gather(enp_type, enp_number, enep:enpp) %>% 
  left_join(lces) %>% 
  distinct() %>% 
  mutate(enp_round = as.numeric(round(enp_number, 1)),
         decade_sd = floor_date(election_date, years(10)))
```

```{r election system elec subset, include = FALSE}
# add subsets of electoral systems to environment (for elections)
e_list <- setNames(split(election_enp, election_enp$v2elloelsy_l), sort(unique(election_enp$v2elloelsy_l)))
list2env(e_list, globalenv())
```

___

## Elections: Effective Number of Parties {.tabset .tabset-fade}

___

The *Effective Number of Parties* (Laakso/Taagepera 1979) is a measurement concept which indicates an adjusted value for a country's party system. 

The value itself can be divided into two measures:

1. ENEP: Effective Number of ***Electoral*** Parties
2. ENPP: Effective Number of ***Parliamentary*** Parties

The *Effective Number of Parties* is calculated as follows: $$ENP = {\frac{1}{\sum_{i = 1}^n p_i^2}}$$

The formula divides 1 by the sum  of each party's relative share $p^2$. The proportion *p* indicates either each party's vote (***ENEP***) or seat (***ENPP***) share.

The values are displayed for each country grouped by region.

___

```{r election plot function, message = FALSE, warning = FALSE}
plot_enp <- function(x) {
  
  subset_e <- election_enp %>% # subset for looping
    filter(country_name == x, 
           election_type == "parliament")
  
  ggplot(subset_e, aes(x = election_date, y = enp_number, group = enp_type, color = enp_type)) +
    geom_line() +
    geom_point() + 
    scale_x_date(breaks = seq(min(subset_e$decade_sd), Sys.Date(), by = "5 years"), 
               date_label = "%Y") +
    scale_y_continuous(limits = c(0, max(subset_e$enp_round)),
                       breaks = seq(0, 15, 1)) +
    scale_color_discrete(labels = c("ENEP", "ENPP")) +
    theme_light() + 
    theme(
      legend.title = element_blank(),
      axis.text.x = element_text(angle = 45, vjust = 0.5, size = 10),
      text = element_text(family = "Formata-CondensedLight", size = 10),
      plot.title = element_text(hjust = 0.5, size = 14),
      axis.text.y = element_text(size = 10)
    ) +
    ggtitle(paste0("\n", subset_e$country_name)) +
    labs(
      x = "Year",
      y = "ENP"
    )
}
```

### All

```{r message = FALSE, warning = FALSE, fig.height = 16, fig.width = 12}
    ggplot(election_enp, aes(x = year, y = enp_number, group = enp_type, color = enp_type)) +
    geom_line() +
    facet_wrap(~ country_name, ncol = 4) +
    scale_x_continuous(breaks = seq(min(election_enp$year), max(election_enp$year) + 5, by = 10)) +
    scale_y_continuous(limits = c(0, max(election_enp$enp_round)),
                       breaks = seq(0, 15, 2)) +
    scale_color_discrete(labels = c("ENEP", "ENPP")) +
    theme_light() + 
    theme(text = element_text(family = "Formata-CondensedLight", size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 30),
      axis.text.x = element_text(angle = 45, vjust = 0.5, size = 10),
      plot.title = element_text(hjust = 0.5, size = 20),
      axis.text.y = element_text(size = 12, face = "bold"),
      panel.spacing = unit(1, "lines")
    ) +
    ggtitle(paste0("\n", "Effective Number of Parties")) +
    labs(
      x = "Year",
      y = "ENP"
    )
```

### First-past-the-post

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`First-past-the-post`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### Alternative vote (SMD)

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Alternative vote (SMD)`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### Parallel (SMD/PR)

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Parallel (SMD/PR)`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### Mixed-member PR

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Mixed-member PR`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### List PR with small MMDs

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`List PR with small MMDs`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### List PR with large MMDs

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`List PR with large MMDs`$country_name)) {
  print(plot_enp(x = country_name))
}
```

### STV

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`STV`$country_name)) {
  print(plot_enp(x = country_name))
}
```