---
title: "ParlGov Visualizations -- Parties"
output:
  html_document:
    code_folding: hide
    fig_width: 8 
    fig_height: 5 
    theme: cosmo
---

```{r package, include = FALSE}
library(tidyverse)
library(countrycode)
library(grid)
library(ggrepel)
```

```{r dataset, include = FALSE}
party_raw <- read_csv("https://parlgov.org/data/parlgov-development_csv-utf-8/view_party.csv", 
                          locale = locale(encoding = "utf-8"))
```

```{r data wrangling and precaution, include = FALSE}
## Party ----
fam_level <- c("com", "eco", "soc", "agr", "lib", "chr", "con", "right")

# countrycode informations
party <- party_raw %>% 
  mutate(region = as.factor(countrycode(country_name, "country.name", "region23")),
         family = factor(family_name_short, fam_level),
         fam_color = case_when(
           family == "com" ~ "#FB9A99",
           family == "eco" ~ "#33A02C",
           family == "soc" ~ "#E31A1C",
           family == "agr" ~ "#B2DF8A",
           family == "lib" ~ "#FDBF6F",
           family == "chr" ~ "#FF7F00",
           family == "con" ~ "#1F78B4",
           family == "right" ~ "#A6CEE3"
         ),
         region_ = (sub(" ", "_", region))) %>% 
  filter(!is.na(family))

# party color
pfcols <- as.character(unique(party$fam_color))
names(pfcols) <- as.character(unique(party$family))

# custom annotations for axis labeling
lr_0 <- textGrob("Extreme left", gp = gpar(fontsize = 9, fontfamily = "Formata"))
lr_10 <- textGrob("Extreme right", gp = gpar(fontsize = 9, fontfamily = "Formata"))
```

```{r region subset, include = FALSE}
# add subsets of region to environment (for parties)
r_list <- setNames(split(party, party$region), sort(unique(party$region)))
list2env(r_list, globalenv())
```
___

## Party Families {.tabset .tabset-fade}

___

The Party Families are grouped by region and each country is plotted severally. The graphs are visualizing the ideological span length of the parties and its families.

___

Each Party Family has a certain abbreviation:

* com -- Communist/Socialist
* eco -- Green/Ecologist
* soc -- Social demoncracy
* agr -- Agrarian
* lib -- Liberal
* chr -- Christian democracy
* con -- Conservative
* right -- Right-wing

___

```{r party plot function, message = FALSE, warning = FALSE}
plot_party_families_lr <- function(x) {
  
  subset_p <- party %>% # subset for looping
    filter(country_name == x)

  ggplot(data = subset_p, 
         aes(x = family, y = left_right, color = family)) + 
    geom_point() +
    geom_line() + 
    coord_flip(clip = "off") +
    scale_color_manual(values = pfcols) +
    scale_y_continuous(breaks = seq(0, 10, by = 1),
                       limits = c(0, 10)) + 
    annotation_custom(lr_0, xmin = -0.25, xmax = -0.25, ymin = 0, ymax = 0) +
    annotation_custom(lr_10, xmin = -0.25, xmax = -0.25, ymin = 10, ymax = 10) + 
    theme_light() + 
    ggtitle(paste0("\n", subset_p$country_name)) +
    labs(
      x = "",
      y = "",
      subtitle = "0 = (Extreme) Left | 10 = (Extreme) Right",
      color = "Party Family"
    )  +
    theme(text = element_text(family = "Formata-CondensedLight", size = 10),
          legend.text = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 14),
          plot.subtitle = element_text(hjust = 0.5, size = 12),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.text.x = element_text(size = 12),
          plot.margin = unit(c(0, 0, 1, 0.5), "cm"))
}
```

### All

```{r message = FALSE, warning = FALSE, fig.height = 16, fig.width = 12}
ggplot(data = subset(party, !is.na(party$family)), 
       aes(x = family, y = left_right, color = family)) + 
    geom_line() + 
    geom_point() +
    facet_wrap(~ country_name, ncol = 4, scales = "free") +
    coord_flip(clip = "off") +
    scale_color_manual(values = pfcols) +
    scale_y_continuous(breaks = seq(0, 10, by = 1),
                       limits = c(0, 10),
                       name = "Left-Right scale") + 
    theme_light() + 
    ggtitle(paste0("\n", "Party Families")) +
    labs(
      x = "",
      y = "",
      color = "Party Family"
    )  +
    theme(text = element_text(family = "Formata-CondensedLight", size = 15),
          legend.title = element_text(size = 20),
          legend.text = element_text(size = 17.5),
          plot.title = element_text(hjust = 0.5, size = 25),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.text.x = element_text(size = 12.5),
          plot.margin = unit(c(0, 0, 1, 0.5), "cm"),
          panel.spacing = unit(1, "lines")
          )
```

### Australia and New Zealand

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Australia and New Zealand`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Eastern Asia

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Eastern Asia`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Eastern Europe

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Eastern Europe`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Northern America

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Northern America`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Northern Europe

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Northern Europe`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Southern Europe

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Southern Europe`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Western Asia

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Western Asia`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```

### Western Europe

```{r message = FALSE, warning = FALSE}
for(country_name in unique(`Western Europe`$country_name)) {
  print(plot_party_families_lr(x = country_name))
}
```